---
title: "Peru Mosquito Analysis"
author: "David Bick"
output: pdf_document
---

#Changes#

* Removed all entries with no species entered
* Instead of color coded graphs, added labels on the axes
* For elevation breakdown by river and species, scales are now consistent over all graphs for easier comparison

```{r  results='hide', message=FALSE, warning=FALSE, echo=F}
library(tidyverse)
library(plyr)
load("peru_data.RData")
peru.df = peru.df %>% filter(Species != "Unlabeled Species")
peru.df = peru.df %>% filter(Taxonomy.Notes != "Unidentified")
```

#Number of Collections Per River#
```{r echo=F, warning=F, message=F}

library(knitr)
library(kableExtra)
library(DT)
peru.df %>%
  group_by(Sector) %>%
  dplyr::summarize(Number=length(unique(Collection.Date))) %>%
  kable() %>% kable_styling()

```

#Number of Village Collection Sites#

```{r echo=F, warning=F, message=F}
peru.df %>%
  group_by(Sector) %>%
  dplyr::summarize(num=length(unique(Exact.Site))) %>%
  kable() %>% kable_styling()
```

#Breakdown of Species By River #  
```{r message=FALSE, warning=FALSE, echo=F}
ggplot(peru.df, aes(x=Species)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Total Collected Species Breakdown") + coord_flip() + labs(y="Count")
ggplot(yavari.df, aes(x=Species)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Yavari River Species Breakdown") + coord_flip() + labs(y="Count")
ggplot(morona.df, aes(x=Species)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Morona River Species Breakdown") + coord_flip() + labs(y="Count")
ggplot(corrientes.df, aes(x=Species)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Corrientes River Species Breakdown") + coord_flip() + labs(y="Count")
ggplot(urituyacu.df, aes(x=Species)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Urituyacu River Species Breakdown") + coord_flip() + labs(y="Count")
```

#Breakdown of Taxonomical Identification By River#  
```{r echo=F}
ggplot(peru.df, aes(x=Taxonomy.Notes)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Total Collected Taxonomical Breakdown") + labs(x="Taxonomical ID", y="Count") + coord_flip()
ggplot(yavari.df, aes(x=Taxonomy.Notes)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Yavari River Taxonomical Breakdown") + labs(x="Taxonomical ID", y="Count") + coord_flip() 
ggplot(morona.df, aes(x=Taxonomy.Notes)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Morona River Taxonomical Breakdown") + labs(x="Taxonomical ID", y="Count") + coord_flip()
ggplot(corrientes.df, aes(x=Taxonomy.Notes)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Corrientes River Taxonomical Breakdown") + labs(x="Taxonomical ID", y="Count") + coord_flip()
ggplot(urituyacu.df, aes(x=Taxonomy.Notes)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle("Urituyacu River Taxonomical Breakdown") + labs(x="Taxonomical ID", y="Count") + coord_flip()
```

#Elevation By River#  
```{r echo=F}
ggplot(elev.plot.df, aes(Elev)) + 
  geom_histogram(binwidth=15) + 
  facet_wrap(~Sector) + labs(x="Elevation", y="Count")
```

#Elevation Breakdown by Species#  
```{r echo=F}
ggplot(elev.plot.df, aes(Elev)) + 
  geom_histogram(binwidth=15) + 
  facet_wrap(~Species) + labs(x="Elevation", y="Count")
```

#Sampling Protocol Breakdown#  
```{r echo=F}

ggplot(peru.df, aes(Sampling.Protocol)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(size=8.5), plot.title=element_text(hjust=0.5), axis.title.y = element_text(size=8.5)) + ylab("# of mosquitos") + xlab("Sampling Protocol") + ggtitle("Breakdown of Sampling Protocol Used") + labs(fill="Sampling Protocol")

```

#Breakdown of Species By Sampling Protocol Used#
```{r echo=F}
breakdown.species = function(df, text) {
  print(ggplot(df, aes(x=Species)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle(text) + coord_flip() + labs(x="Species", y="Count"))
}

d_ply(peru.df, .(Sampling.Protocol), function(x) breakdown.species(x, x$Sampling.Protocol[1]))
```

#Breakdown of Taxonomical Identification By Sampling Protocol Used#
```{r echo=F}
breakdown.taxonomy = function(df, text) {
  print(ggplot(df, aes(x=Taxonomy.Notes)) + 
  geom_bar() + 
  theme(axis.text.x = element_text(), plot.title=element_text(hjust=0.5), axis.title.x = element_text()) + 
  ggtitle(text) + labs(x="Taxonomical ID", y="Count") + coord_flip())
}

d_ply(peru.df, .(Sampling.Protocol), function(x) breakdown.taxonomy(x, x$Sampling.Protocol[1]))
```













