---
title: "Mali Project"
author: "David Bick"
date: "7/11/2018"
output: pdf_document
---

## Changes ##

* Removed all entries with undocumented species or month - this happened to remove all observations from April
* From this, the sex breakdown by altitude changed, with males being more common at higher altitude
* Added movement of perexiguus and anopheles by months andthroughout the years
* Easier to see that Culex mali sp. 1 proportion increases as altitude increases

```{r echo=F, warning=F}
suppressMessages(library(ggthemes))
suppressMessages(library(tidyverse))
suppressMessages(library(ggplot2))
suppressMessages(library(dplyr))
suppressMessages(library(knitr))
suppressMessages(library(kableExtra))
suppressMessages(library(ggpubr))
mali.df = read.csv("mali.csv", stringsAsFactors = FALSE)
mali.df$altitude = mali.df$Collection.Notes %>% substr(., nchar(.)-3, nchar(.)-1) %>% trimws() %>% as.numeric()
get.month = function(x) {
  if (x == "") return ("Unlabeled")
  return(strsplit(x, "-") %>% "[["(., 1) %>% "[["(., 2))
}
get.year = function(x) {
  if (x == "") return ("Unlabeled")
  return(strsplit(x, "-") %>% "[["(., 1) %>% "[["(., 3))
}
mali.df$year = lapply(mali.df$Collection.Date, get.year) %>% unlist()
mali.df$month = lapply(mali.df$Collection.Date, get.month) %>% unlist()
mali.df$Species[mali.df$Species == ""] = "Unlabeled"
mali.df = mali.df[mali.df$Species != "Unlabeled",]
mali.df = mali.df[mali.df$month != "Unlabeled",]
mali.df$posix.date = mali.df$Collection.Date %>% strptime(., format="%d-%b-%y") %>% as.POSIXct()
mali.df$year.month = mali.df$posix.date %>% format(., format="%y-%m")

```



## Perexiguus Over The Year##  
```{r echo=F, warning=F}
months = c("May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

perexiguus = mali.df %>% filter(Species == "Culex perexiguus")
one = perexiguus %>% 
  group_by(month) %>%
  count()
one$month = factor(one$month, levels=months)
one = one[c(5,4,3,1,8,7,6,2),]
one$Percent = ((one$n / sum(one$n))*100) %>% round(., 2)
labels = paste(one$Percent, "%", sep="")

ggplot(one, aes(x=month, y=n)) + geom_bar(stat="identity") + 
  labs(y="# of mosquitos", x="Month", title="Number of Culex Perexiguus By Month") + 
  theme(plot.title = element_text(hjust=0.5)) + 
  geom_text(aes(label=labels), vjust=-0.4)

two = perexiguus[-(c(1, 2, 3)),]  %>%
  group_by(year.month) %>%
  count()

years = c("13", "14", "15")
months = c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
year.months = c()
for (year in years) {
  for (month in months) {
    new = paste(year, month, sep="-")
    year.months=c(year.months, new)
  }
}

two = two[-c(1,2),]
full = data.frame(year.month=year.months, n=vector(length=length(year.months), mode="numeric"), stringsAsFactors = FALSE)

for (year.month in two$year.month) {
  full[full$year.month == year.month,]$n = two[two$year.month == year.month,]$n
}

ggplot(full, aes(x=year.month, y=n)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(size=7, angle=90), plot.title = element_text(hjust=0.5)) + 
  labs(x="Date: YY-MM ", y="# of mosquitos", title="Perexiguus by Month 2013-2015")
```

## Anopheles Over The Year##  
```{r echo=F}
months = c("May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

anopheles = mali.df %>% filter(Genus == "Anopheles")
one = anopheles %>% 
  group_by(month) %>%
  count()
one$month = factor(one$month, levels=months)
one = one[c(5,4,3,1,8,7,6,2),]
one$Percent = ((one$n / sum(one$n))*100) %>% round(., 2)
labels = paste(one$Percent, "%", sep="")

ggplot(one, aes(x=month, y=n)) + geom_bar(stat="identity") + 
  labs(y="# of mosquitos", x="Month", title="Number of Anopheles By Month") + 
  theme(plot.title = element_text(hjust=0.5)) + 
  geom_text(aes(label=labels), vjust=-0.4)

two = anopheles[-(c(1, 2, 3)),]  %>%
  group_by(year.month) %>%
  count()

years = c("13", "14", "15")
months = c("01", "02", "03", "04", "05", "06", "07", "08", "09", "10", "11", "12")
year.months = c()
for (year in years) {
  for (month in months) {
    new = paste(year, month, sep="-")
    year.months=c(year.months, new)
  }
}

two = two[-c(1,2),]
full = data.frame(year.month=year.months, n=vector(length=length(year.months), mode="numeric"), stringsAsFactors = FALSE)

for (year.month in two$year.month) {
  full[full$year.month == year.month,]$n = two[two$year.month == year.month,]$n
}

ggplot(full, aes(x=year.month, y=n)) + geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(size=7, angle=90), plot.title = element_text(hjust=0.5)) + 
  labs(x="Date: YY-MM ", y="# of mosquitos", title="Anopheles by Month 2013-2015")
```

##Counts of Mosquitos at Each Altitude##
```{r echo=F, warning=F}
alt.counts = mali.df %>%
  group_by(altitude) %>% 
  summarize(Count=n()) %>%
  arrange(desc(altitude))
alt.counts = alt.counts[!is.na(alt.counts$altitude),]
alt.counts %>% kable() %>% kable_styling(bootstrap_options = c("condensed"))
```

##Overall Species Breakdown Top 10 Most Common ##
```{r echo=F}
overall.species.breakdown = mali.df %>% 
  group_by(Species) %>%
  summarize(Count=n()) %>%
  arrange(desc(Count))
overall.species.breakdown$Percent = ((overall.species.breakdown$Count / sum(overall.species.breakdown$Count)) * 100) %>% round(., 2)
head(overall.species.breakdown, 10) %>% kable() %>% kable_styling()
 
```

##Anopheles by Altitude##
```{r echo=F}

anoph.df = mali.df %>% filter(Genus == "Anopheles")
anoph = anoph.df %>%
  group_by(altitude) %>%
  summarize(Count=n()) %>%
  arrange(desc(altitude))
anoph$Genus = "Anopheles"
anoph$Percent = ((anoph$Count / sum(anoph$Count)) * 100) %>% round(., 2)
anoph = anoph[!is.na(anoph$altitude),]
anoph = anoph[,c(3, 1, 2, 4)]
#anoph %>% kable() %>% kable_styling()

labels = paste(anoph$Percent, "%", sep="")
ggplot(anoph, aes(x=reorder(altitude, altitude), y=Count)) + 
  geom_bar(stat="identity") + labs(x="Altitude") + ggtitle("Percent of Total Anopheles at Each Altitude") + 
  theme(plot.title = element_text(hjust=0.5)) + geom_text(aes(label=labels), vjust=-0.4, color="black")

```


##Sex of Mosquito Over Altitude##
```{r echo=F}

library(ggpubr)
full.values = mali.df[mali.df$Sex != "",]
full.values = mali.df[!is.na(mali.df$altitude),]
full.values.f = full.values %>% filter(Sex == "F")
full.values.m = full.values %>% filter(Sex == "M")

one = ggplot(full.values.f, aes(reorder(altitude, altitude))) + 
  geom_bar() + ggtitle("Female") + theme(plot.title=element_text(hjust=0.5)) + 
  labs(x="Altitude")

two = ggplot(full.values.m, aes(reorder(altitude, altitude))) + 
  geom_bar() + ggtitle("Male") + theme(plot.title=element_text(hjust=0.5)) + 
  labs(x="Altitude")

ggarrange(one, two)


```

##Altitudinal Species Breakdown with 100 meter Aggregations##  
```{r echo=F}

species.breakdown.alt.group.100 = function(alt, viz) {
  df = mali.df %>% filter(alt.group.100 == alt)
  table = df %>% group_by(Species) %>% summarize(Count=n()) %>% arrange(desc(Count))
  table$Altitude.Range = alt
  table = table[,c(3, 1, 2)]
  table$Percent = ((table$Count / sum(table$Count)) * 100) %>% round(., 2)
  small.table = table[table$Percent >= 3,]
  labels = paste(small.table$Percent, "%", sep="")
  plot = ggplot(small.table, aes(x=reorder(Species, Count), y=Count)) + geom_bar(stat="identity") + theme_tufte()
  plot = plot + geom_text(aes(label=labels), hjust=1.1, color="white", size=2.4)
  plot = plot + theme(plot.title = element_text(hjust=0.5), 
                      axis.text.y = element_text(size=9)) + coord_flip() + 
    ggtitle(paste("Altitude Range", alt)) + labs(x="Species")
  if (viz) print(plot)
  return(list(plot=plot,
              table=table))
}

mali.df$alt.group.100 = cut(mali.df$altitude, breaks=c(0, 100, 200, 300))
alt.groups = mali.df$alt.group.100 %>% unique %>% sort()
species.breakdowns.alt.groups.100 = lapply(alt.groups, function(alt) species.breakdown.alt.group.100(alt, T))
#plots = (lapply(species.breakdowns.alt.groups.100, function(x) "[["(x, 1)))
#figure = ggarrange(plots[[1]], plots[[2]], plots[[3]])
#figure

```

##Altitudinal Species Breakdown with 50 meter Aggregations##  

250-300 range differs most from the rest, with culex mali sp.1 replacing culex perexiguus at the top  
An. pharoensis found in relatively high quantity only at the top altitude range  
```{r echo=F}

par(mfrow=c(2,3))
species.breakdown.alt.group.50 = function(alt, viz) {
  df = mali.df %>% filter(alt.group.50 == alt)
  table = df %>% group_by(Species) %>% summarize(Count=n()) %>% arrange(desc(Count))
  table$Altitude.Range = alt
  table = table[,c(3, 1, 2)]
  table$Percent = ((table$Count / sum(table$Count)) * 100) %>% round(., 2)
  small.table = table[table$Percent >= 3,]
  labels = paste(small.table$Percent, "%", sep="")
  plot = ggplot(small.table, aes(x=reorder(Species, Count), y=Count)) + geom_bar(stat="identity") + theme_tufte()
  plot = plot + geom_text(aes(label=labels), hjust=1.1, color="white", size=3)
  plot = plot + theme(plot.title = element_text(hjust=0.5), 
                      axis.text.y = element_text(size=9)) + coord_flip() + 
    ggtitle(paste("Altitude Range", alt)) + labs(x="Species")
  if (viz) print(plot)
  return(list(plot=plot,
              table=table))
}


mali.df$alt.group.50 = cut(mali.df$altitude, breaks=c(0, 100, 150, 200, 250, 300))
alt.groups = mali.df$alt.group.50 %>% unique %>% sort()
species.breakdowns.alt.groups.50 = lapply(alt.groups, function(alt) species.breakdown.alt.group.50(alt, T))
plots.1 = (lapply(species.breakdowns.alt.groups.50, function(x) "[["(x, 1)))


```

##Species Breakdown by Month ## 

  *Culex perexiguus becomes the most common species in May and keeps the top spot until November  
  *November is the first month where anopheles are the most common   
```{r echo=F}

months = c("Mar", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")

species.breakdown.month = function(mon, viz) {
  df = mali.df %>% filter(month == mon)
  table = df %>% group_by(Species) %>% summarize(Count=n()) %>% arrange(desc(Count))
  table$month = mon
  table = table[,c(3, 1, 2)]
  table$Percent = ((table$Count / sum(table$Count)) * 100) %>% round(., 2)
  small.table = table[table$Percent >= 3,]
  labels = paste(small.table$Percent, "%", sep="")
  plot = ggplot(small.table, aes(x=reorder(Species, Count), y=Count)) + geom_bar(stat="identity") + theme_tufte()
  plot = plot + geom_text(aes(label=labels), hjust=1.1, color="white", size=3)
  plot = plot + theme(plot.title = element_text(hjust=0.5), 
                      axis.text.y = element_text(size=9)) + coord_flip() + 
    ggtitle(paste("Month:", toupper(mon))) + labs(x="Species")
  if (viz) print(plot)
  return(list(plot=plot,
              table=table))
}

month.breakdowns = lapply(months, function(mon) species.breakdown.month(mon, T))



```



